\chapter*{Acknowledgements}\label{ch:ac} % no number
\addcontentsline{toc}{chapter}{Acknowledgements}% put it in the ToC

It would be far too lengthy to acknowledge everyone who contributed to this work, and the many ways they did so. I will therefore name a few people, in no particular order, and mention at least one thing I consider relevant about them, even if it is not necessarily the most important.

I should thank all my friends around the world, but a few special mentions are in order. The first thing Michele told me when we met was that I should never have gone into epistemic game theory; I did my best anyway. Consistently, one of the last things he told me was that a paper in this thesis will probably never be published in a finance journal. To Pau I owe inspiration for a new definition of certainty: an event as likely as finding him in the office when I arrive. My friends from the Caba supported me with words consisting only of one capital letter, but full of meaning. I thank my brothers, to whom this thesis is dedicated, for always playing fair with me even when I lost. I hope we keep learning to play together. I owe gratitude to many professors, but I will mention only a few. François never tired of telling me that my notation was terrible. Interestingly, he once asked about the connection between Bayes rule and mixing liquids. The first time I met Marciano he called one of my ideas a \textquote{bel trucco}. I remain indebted to him for giving me the courage and the support to develop it into a chapter of this thesis. Ingela not only took a bet on me against the odds, she also encouraged me to pursue original ideas and showed me how to nurture them. She praised the long-gone quotations that once opened these chapters—quotes I never had the courage to keep.

Ringrazio i miei genitori, che hanno sempre intravisto per me un futuro da insegnante e ricercatore, sopportando con pazienza i molti alti e bassi del percorso. Mio padre mi ha insegnato a guardare in alto anziché fissare i miei piedi; ho seguito quel consiglio con tale entusiasmo da passare fin troppo tempo a contemplare il cielo. Per fortuna mia madre mi ha sempre ricordato di camminare con i piedi ben dritti. Da loro ho imparato a fantasticare con ordine, un equilibrio ossimorico solo in apparenza che spero si rifletta in questa tesi.

Across continents, for more than a third of my life, and throughout my entire intellectual journey, I have been in Annalisa's loving companionship. We bonded over a shared urge to understand our choices and give them meaning, so much so that it is often hard to tell whether we are discussing research or ourselves. She sees how deeply this thesis reflects my ideas, methods, and aspirations, yet kindly forgives my prose for not matching her eloquence. I thank her for her unique ability to interpret every expression of myself, usually over a good cappuccino in a \textquote{suggestive} spot.